The Near and Mid-infrared photometric properties of known redshift $z\geq5$ Quasars

### 2019-06-14 tex/VHzQ.tex
So friggin nearly there...

\maketitle
\begin{abstract}
We assemble a catalogue of 463 spectroscopically confirmed very high
($z\geq5.00$) redshift quasars and report their near- ($ZYJHK_{s}$ and
$K$) and mid- (WISE) infrared properties. We find that SDSS and
Pan-STARRS1 together identified over half of the VH$z$Q sample and
that 97.0\% of the VH$z$Q sample is detected in one or more NIR
($ZYJHK/K_{s}$) band, with lack of coverage rather than lack of depth
for the objects that are not detected in the NIR. 362 (78.2\%) VH$z$Qs
are detected at 3.4$\mu$m in the W1 band from the unWISE catalog and
all of the $z\geq7$ quasars are detected in both unWISE W1 and
W2. Using archival WFCAM/UKIRT and VIRCAM/VISTA data we check for
photometric variability in the near-infrared that might be expected
from super-Eddington accretion. We find 32 of the quasars have
sufficient NIR measurements and signal-to-noise to look for
variability. Weak variability was detected in multiple bands of SDSS
J0959+0227, and very marginally in the $Y$-band of MMT J0215-0529. Two
other quasars, SDSS J0349+0034 and SDSS J2220-0101 had significant
differences between their WFCAM and VISTA magnitudes in one band. 
With supermassive black hole accretion likely to be redshift invariant 
up to very high-redshift, further monitoring of these sources is warranted. 
All the data, analysis codes and plots used and generated here can be
found at: \href{https://github.com/d80b2t/VHzQ}{\tt
github.com/d80b2t/VHzQ}.
\end{abstract}

@@

\citet{Secrest2015} present an all-sky sample of $\approx$1.35 million
AGN meeting a two-color, (W1-W2) vs. (W2-W3), photometric selection as
applied to sources from AllWISE. This colour-selection is motivated by
\citet{Mateos2012}. We investigate if this catalogue, which is tuned
for, and noted to be highly complete at, lower, $z\lesssim2$
redshifts, contain the VH$z$Qs. We find only 5 VH$z$Qs from our sample
match the Secrest catalogue, but given a W3 detection is required,
this is not surprising. Further, in addition to the relative lack of
sensitivity in W3, the mid-IR colors are moving out of the
\citet{Mateos2012} color criterion at high redshift. This was shown
using an AGN SED where AGNs above a redshift of about $\sim$2 fall out
of the colour criterion, see e.g. Figure 5 of \citet[][]{Mateos2012}
and Figure 9 of \citet{Secrest2015}.
%%
\citet{Assef2018} released two large catalogues of AGN
candidates identified across 30,000 deg$^2$ of extragalactic sky from
the WISE AllWISE Data Release. The ``R90'' catalogue, contains 4.5M
AGN candidates at 90\% reliability (and $\approx$150 AGN candidates
per deg$^2$) while the ``C75'' catalogue consists of 20.9M AGN
candidates at 75\% completeness (and ($\approx$700 AGN candidates per
deg$^2$). Crossmatching our catalogue of 463 VH$z$Qs with these
catalogues, produces 42 matches with the R90 sample and 98 matches
with the C75 sample. Both catalogues unsurprisingly match to the
ultraluminous quasar SDSS J0100+2802 \citep{Wu2015} while the C75, but
not the R90 catalogue matches to the first redshift $z>7$ quasar ULAS
J1120+0641 \citep{Mortlock2011}. Neither catalogue matches to the
current highest-redshift object J1342+0928 \citep{Banados2018}.

\begin{figure}
    \centering
    \includegraphics[width=8.6cm,  clip, trim=6mm 4mm 6mm 10mm]
    {../color_color/W1W2_plots/W1W2W3/W1W2W3_hexplots_20190610v2.png} 
  \vspace{-14pt}
    \caption{The (W2-W3) vs. (W1-W2) colour – colour diagram showing the WISE colours 
      for the 283 VH$z$Q that have reported w3mpro values (blue to red coloured points).
      400,000 quasars at redshift $z\lesssim3.0$ from the SDSS DR14 quasar catalogue 
      \citep{Paris2018} are also shown
      (grey colour-scale). }
    \label{fig:W1W2W3}
\end{figure}
\subsection{MIR Colours}
Due to the depth all-sky coverage of the WISE (and NEOWISE-(R))
mission, several investigations have looked at how WISE detects AGN
\citep[e.g][]{Stern2012, Assef2012, Secrest2015, LaMassa2017,
Assef2018, Glikman2018, Hviding2018, LaMassa2019}

@@

\begin{figure*}
    \centering
    \includegraphics[width=0.24\textwidth]{../data/WISE/image_thumbnails/UHS_J0439+1634.png} 
    \includegraphics[width=0.24\textwidth]{../data/WISE/image_thumbnails/SDSS_J1443+3623.png}
    \includegraphics[width=0.24\textwidth]{../data/WISE/image_thumbnails/SDSS_J1623+4705.png}
    \includegraphics[width=0.24\textwidth]{../data/WISE/image_thumbnails/SDSS_J0100+2802.png}
    \caption{Four of the brightest MIR VH$z$Qs. 
      The thumbnails are the RGB colour outputs using the AllWISE W1W2W3 bands from the 
      IRSA WISE Image Service, with a scale of 120'' on the side. 
      {\it (a)} UHS J0439+1634 was discovered by
      \citet{Fan2019} and is a strongly lensed quasar at $z=$6.51. 
      {\it (b)} SDSS J0100+2802 is the \citet{Wu2015} object. 
      {\it (c)} SDSS J1443+3623 has {\tt w3mpro, w4mpro } = ($11.189\pm0.1$, $8.699\pm0.26$) and 
      {\it (d)} SDSS J1623+4705 has {\tt w3mpro, w4mpro } = ($11.425\pm0.1$, $8.954\pm0.28$).}
    \label{fig:VHzQ_W4s}
\end{figure*}
The VH$z$Q with the bluest colour in (W1-W2) is DELS J1048-0109 with (W1-W2) = 0.154. 
and the reddest object in (W1-W2) is SHELLQs J2223+0326  with (W1-W2) =  1.33.

The set of VH$z$Q detected in WISE W3 and W4 contains 51 objects that are
detected in the broad W3 filter and 10 objects that are formally detected
in W4 ({\tt w4snr}$\geq$3.0). 
%%
Four of the W3/4 objects are presented in \label{fig:VHzQ_W4s}. 
UHS J0439+1634 (Fig.~\ref{fig:VHzQ_W4s} (a)) with W4=7.165$\pm$0.12 was discovered by
\citet{Fan2019} and is a strongly lensed quasar at $z=$6.51. This high
luminosity is mostly not intrinsic, but is boosted by an intervening
redshift $z\sim$0.7 galaxy. 
Other high-redshift quasars that are bright W3/4 objects may well 
also so be lensed \citep[e.g.,][]{Glikman2018lens, Fan2019}, 
but high-resolution follow-up is needed to confirm. 
%%
SDSS J0100+2802 (redshift $z=6.33$) does
not have a formal W4 detection (W4 $=8.98\pm0.45$ and {\tt w4snr}=2.4),
but as reported by \citet{Wu2015}, has a bright detection in the
2MASS $J$, $H$ and $K_{\rm s}$ bands and is also the bluest object 
in both (W1-W3) and (W2-W3) with values of 2.73 and 1.93, respectively, 
so we report it here.
%%
PSO J075.9356-07.5061 is the reddest W3 detected object with (W1-W3) =
5.31 and (W2-W3)= 4.65.

>>>>

\maketitle
\begin{abstract}
We assemble a catalogue of 463 spectroscopically confirmed very high
($z\geq5.00$) redshift quasars and report their near- ($ZYJHK_{s}/K$)
and mid- (WISE W1234) infrared properties. 97\% of the VH$z$Q sample
is detected in one or more NIR ($ZYJHK/K_{s}$) band, with lack of
coverage rather than lack of depth being the reason for the
non-detections. 362 (78\%) of the very high redshift quasars are
detected at 3.4$\mu$m in the W1 band from the unWISE catalog and all
of the $z\geq7$ quasars are detected in both unWISE W1 and W2. Using
archival WFCAM/UKIRT and VIRCAM/VISTA data we check for photometric
variability that might be expected from super-Eddington accretion. We
find 32 of the quasars have sufficient NIR measurements and
signal-to-noise to look for variability. Weak variability was detected
in multiple bands of SDSS J0959+0227, and very marginally in the
$Y$-band of MMT J0215-0529. Two other quasars, SDSS J0349+0034 and
SDSS J2220-0101 had significant differences between their WFCAM and
VISTA magnitudes in one band.  With supermassive black hole accretion
likely to be redshift invariant up to very high-redshift, further
monitoring of these sources is warranted.  All the data, analysis
codes and plots used and generated here can be found at:
\href{https://github.com/d80b2t/VHzQ}{\tt github.com/d80b2t/VHzQ}.
\end{abstract}

@@

\begin{figure}
    \centering
    \includegraphics[width=8.6cm,  clip, trim=6mm 4mm 6mm 10mm]
    {../color_color/W1W2_plots/W1W2W3/W1W2W3_hexplots_20190610v2.png} 
  \vspace{-14pt}
    \caption{The (W2-W3) vs. (W1-W2) colour – colour diagram showing the WISE colours 
      for the 283 VH$z$Q that have reported w3mpro values (blue to red coloured points).
      400,000 quasars at redshift $z\lesssim3.0$ from the SDSS DR14 quasar catalogue 
      \citep{Paris2018} are also shown
      (grey colour-scale). }
    \label{fig:W1W2W3}
\end{figure}

\citet{Secrest2015} present an all-sky sample of $\approx$1.35 million
AGN meeting a two-color, (W1-W2) vs. (W2-W3), photometric selection as
applied to sources from AllWISE. This colour-selection is motivated by
\citet{Mateos2012}. We investigate if this catalogue, which is tuned
for, and noted to be highly complete at, lower, $z\lesssim2$
redshifts, contain the VH$z$Qs. We find only 5 VH$z$Qs from our sample
match the Secrest catalogue, but given a W3 detection is required,
this is not surprising. Further, in addition to the relative lack of
sensitivity in W3, the mid-IR colors are moving out of the
\citet{Mateos2012} color criterion at high redshift. This was shown
using an AGN SED where AGNs above a redshift of about $\sim$2 fall out
of the colour criterion, see e.g. Figure 5 of \citet[][]{Mateos2012}
and Figure 9 of \citet{Secrest2015}.
%%
\citet{Assef2018} released two large catalogues of AGN
candidates identified across 30,000 deg$^2$ of extragalactic sky from
the WISE AllWISE Data Release. The ``R90'' catalogue, contains 4.5M
AGN candidates at 90\% reliability (and $\approx$150 AGN candidates
per deg$^2$) while the ``C75'' catalogue consists of 20.9M AGN
candidates at 75\% completeness (and $\approx$700 AGN candidates per
deg$^2$). Crossmatching our catalogue of 463 VH$z$Qs with these
catalogues, produces 42 matches with the R90 sample and 98 matches
with the C75 sample. Both catalogues unsurprisingly match to the
ultraluminous quasar SDSS J0100+2802 \citep{Wu2015} while the C75, but
not the R90 catalogue matches to the first redshift $z>7$ quasar ULAS
J1120+0641 \citep{Mortlock2011}. Neither the R90 or C75 catalogue matches to the
current highest-redshift object J1342+0928 \citep{Banados2018}.

\begin{figure*}
    \centering
    \includegraphics[width=0.24\textwidth]{../data/WISE/image_thumbnails/DELS_J1048-0109.png} 
    \includegraphics[width=0.24\textwidth]{../data/WISE/image_thumbnails/UHS_J0439+1634.png} 
    \includegraphics[width=0.24\textwidth]{../data/WISE/image_thumbnails/SDWISE_J0306+1853.png}
    \includegraphics[width=0.24\textwidth]{../data/WISE/image_thumbnails/SDSS_J0100+2802.png}
    \caption{VH$z$Qs with interesting MIR properties. 
      The thumbnails are the RGB colour outputs using the AllWISE W1W2W3 bands from the 
      IRSA WISE Image Service, with a scale of 120'' on the side. 
      {\it (a)} DELS J1048-0109 is the bluest in (W1-W2) colour quasar. 
      {\it (b)} UHS J0439+1634 was discovered by \citet{Fan2019} and is a strongly lensed quasar at $z=6.51$. 
      {\it (c)} SDWISE J0306+1853 has {\tt w3mpro} = $10.51\pm0.1$ and {\tt w4mpro } = $7.586\pm0.16$. 
      {\it (d)} SDSS J0100+2802 is the \citet{Wu2015} object.  
%      {\it (d)} SDSS J1443+3623 has {\tt w3mpro, w4mpro } = ($11.189\pm0.1$, $8.699\pm0.26$) and 
  %    {\it (d)} SDSS J1623+4705 has {\tt w3mpro, w4mpro } = ($11.425\pm0.1$, $8.954\pm0.28$).
}
    \label{fig:VHzQ_W4s}
\end{figure*}
\subsection{MIR Colours}
Due to the depth all-sky coverage of the WISE (and NEOWISE-(R))
mission, several investigations have looked at how WISE detects AGN
\citep[e.g][]{Stern2012, Assef2012, Secrest2015, LaMassa2017,
Assef2018, Glikman2018, Hviding2018, LaMassa2019}

@@

The VH$z$Q with the bluest colour in (W1-W2) is DELS~J1048-0109, with (W1-W2) = 0.154, and the WISE W1W2W3 color image is shown in Fig.~\ref{fig:VHzQ_W4s} {\it(a)}. The reddest object in (W1-W2) is SHELLQs~J2223+0326 with (W1-W2) = 1.33.

The set of VH$z$Q detected in WISE W3 and W4 contains 51 objects that are detected in the broad W3 filter and 10 objects that are formally detected in W4 ({\tt w4snr}$\geq$3.0). Three of the W3/4 objects are presented in Figure~\ref{fig:VHzQ_W4s}. UHS~J0439+1634 (Fig.~\ref{fig:VHzQ_W4s} {\it (b)}) with W4=7.165$\pm$0.12 was discovered by \citet{Fan2019} and is a strongly lensed quasar at $z=$6.51. This high luminosity is mostly not intrinsic, but is boosted by an intervening redshift $z\sim$0.7 galaxy. Other high-redshift quasars that are bright W3/4 objects e.g. SDWISE~J0306+1853 (Fig.~\ref{fig:VHzQ_W4s} {\it(c)}) may well also so be lensed \citep[e.g.,][]{Glikman2018lens, Fan2019}, but high-resolution follow-up is needed to confirm. SDSS~J0100+2802 (redshift $z=6.33$) does not have a formal W4 detection (W4 $=8.98\pm0.45$ and {\tt w4snr}=2.4), but as reported by \citet{Wu2015}, has a detection in the 2MASS $J$, $H$ and $K_{\rm s}$ bands and is also the bluest object in both (W1-W3) and (W2-W3) with values of 2.73 and 1.93, respectively, and is shown in Fig.~\ref{fig:VHzQ_W4s} {\it(d)}.
